---
output:
  md_document:
    variant: markdown_github
---

# Purpose

This work folder was created to be used in the evaluation for the Financial Econometrics course presented at Stellebosch University for the purpose of completing my Mcom (Economics). 
It outlines ... and for this practical, I will attempt to do ... 

```{r, include=FALSE}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
```

Below, the r setup used throughout. Importantly, the packages loaded for each question vary depending on the respective requirements.  



```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, ig.align = "left", fig.height = 3, fig.pos = "H", fig.width = 5,
                      message=FALSE, warning=FALSE, comment = NA)
library(pacman)
pacman::p_load(fmxdat, Texevier, knitr, kableExtra, tidyverse)

```



# Description

This folder was created using the fmxdat package (https://github.com/Nicktz/fmxdat) with the following code:
```{r, eval=FALSE}
fmxdat::make_project()
```

In addition, the folders used for the seperate questions (see *Question 1* and *Question 2*) were created using the Texevier package (https://github.com/Nicktz/Texevier):

```{r, eval=FALSE}

Texevier::create_template_html(directory = 'Questions', template_name = 'Question_1')
# Alternatively, for knitting to PDF: Texevier::create_template(directory = 'Questions', template_name = 'Question_1')

```


## Data

The relevant data used in completing this practical was sourced from: https://www.fmx.nfkatzke.com/FMX_data_2021.zip

*PROVIDE SHORT DESCRIPTION OF THE DATA* 











