---
output:
  md_document:
    variant: markdown_github
---

# Purpose

This work folder was created to be used in the evaluation for the Financial Econometrics course presented at Stellebosch University for the purpose of completing my Mcom (Economics). 
It outlines ... and for this practical, I will attempt to do ... 

```{r, include=FALSE}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
```

Below, the r setup used throughout. Importantly, the packages loaded for each question vary depending on the respective requirements.  



```{r setup, message=FALSE}
knitr::opts_chunk$set(fig.align = "left", fig.height = 3, fig.pos = "H", fig.width = 5,
                      message=FALSE, warning=FALSE, comment = NA)
library(pacman)
pacman::p_load(fmxdat, Texevier, knitr, kableExtra, tidyverse, readr)

```



# Description

This folder was created using the fmxdat package (https://github.com/Nicktz/fmxdat) with the following code:
```{r, eval=FALSE}
fmxdat::make_project()
```

In addition, the folders used for the seperate questions (inside the "root" *Questions* folder) were created using the Texevier package (https://github.com/Nicktz/Texevier):

```{r, eval=FALSE}

Texevier::create_template_html(directory = 'Questions', template_name = 'Question_1')
Texevier::create_template_html(directory='Questions', template_name = 'Question_2')


```


# Data

The relevant data used in completing this practical was sourced from: https://www.fmx.nfkatzke.com/FMX_data_2021.zip.

The data sourced was all in .rds format, with accompanying txt files providing brief descriptions for the relevant datasets. Data used to complete this practical was placed in the "data" folder. These datasets, however, are not available on this github (due to storage constraints), and were excluded from the commits by updating the .gitignore file: specifying that objects in "data/" should be ignored. In this folder, there is a README.md file which provides an overview of the data. The data used to complete questions will be specified when these questions are addressed.

Data in .rds format was uploaded to the environment using code in the following format:

```{r, eval=FALSE}

somedat <- read_rds("data/some_data.rds")

```

# Question 1


# Question 2

```{r}
```









